---
title: "RomanEstrada_ASV_analysis"
author: "David Haak"
date: "`r Sys.Date()`"
output: html_document
keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = FALSE,
                      cache = TRUE,
                      echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("/Volumes/GoogleDrive/My Drive/2022 Wheat MS ASVs/Wheat_seed_endophytes/Data/"))
```

```{r}
library(dada2)
library(phyloseq)
library(DESeq2)
library(ggplot2)
library(ape)
library(plyr)
library(dplyr)
library(microeco)
library(mecodev)
library(gridExtra)
library(file2meco)
library(ggh4x)
library(ggdendro)
library(ggpubr)
library(RColorBrewer)
library(ggalluvial)
library(NatParksPalettes)
library(biomformat)
library(lme4)
library(tidyverse)
```

## Convert biom file to csv
```{r}
asv_table <- read_biom("LB5353_wheat_FINAL_TABLE/feature-table.biom")
asv_table <- as.data.frame(as.matrix(biom_data(asv_table)))
view(asv_table)
write.csv(asv_table, "LB5353_wheat_FINAL_TABLE/feature-table.csv")
```

```{r}
asv_table <- read.csv("LB5353_wheat_FINAL_TABLE/feature-table.csv")
tax_file <- read.csv ("LB5353_wheat_FINAL_TAXONOMY/taxonomy.csv")
full_table <- inner_join(asv_table,tax_file, by = "Feature_ID")
full_table$asvid <- paste0("ASV", 1:nrow(full_table))
asv_table <- full_table %>% 
  select(2:28, 35)
tax_tab <- full_table %>% 
  select(29:35)
row.names(asv_table) <- asv_table$asvid
asv_table <- asv_table %>% select(-(asvid))
row.names(tax_tab) <- tax_tab$asvid
tax_tab <- tax_tab %>% 
  select(-asvid) %>% 
  separate(Kingdom, c("a", "Kingdom"), "\\__*") %>% 
  separate(Phylum, c("a", "Phylum"), "\\__*") %>% 
  separate(Class, c("a", "Class"), "\\__*") %>% 
  separate(Order, c("a", "Order"), "\\__*") %>% 
  separate(Family, c("a", "Family"), "\\__*") %>% 
  separate(Genus, c("a", "Genus"), "\\__*") %>% 
  select(-a)

  

kernel_asv <- microtable$new(otu_table = asv_table, tax_table = tax_tab)
kernel_asv$sample_table$Site <- sample_file$WheatSite
kernel_asv$sample_table$Variety <- sample_file$WheatVariety
kernel_asv$sample_table$Rep <- sample_file$WheatSampleSet
kernel_asv$tidy_dataset()
kernel_asv

```

```{r RAplotBySite}
kernel_abund<-trans_abund$new(dataset = kernel_asv, taxrank = "Family", ntaxa = 20)
p1 <- kernel_abund$plot_bar(others_color = "grey70", facet = "Site", xtext_keep = FALSE, legend_text_italic = FALSE, color_palette_20)

p2 <- kernel_abund$plot_bar(others_color = "grey70", facet = "Variety", facet2 = "Site", xtext_keep = FALSE, legend_text_italic = FALSE, barwidth = 1, color_palette_20)
```

```{r Most_abundant}
kernel_abund2 <- trans_abund$new(dataset = kernel_asv, taxrank = "Family", ntaxa = 7)
p3 <- kernel_abund2$plot_box(group = "Site") 
```

## Lineplot of taxa RA by site
```{r lineplotRA}
taxaline <- trans_abund$new(dataset = kernel_asv, taxrank = "Genus", ntaxa = 5)
p4 <- taxaline$plot_line(color_palette_20)
taxalinesite <- trans_abund$new(dataset = kernel_asv, taxrank = "Genus", ntaxa = 8, group = "Site")
p5 <- taxalinesite$plot_line(position = position_dodge(0.3), xtext_type_hor = TRUE, color_palette_20) + theme_bw()

taxalinesite <- trans_abund$new(dataset = kernel_asv, taxrank = "Genus", ntaxa = 6, group = "Variety")
p5.1 <- taxalinesite$plot_line(position = position_dodge(0.3), xtext_type_hor = TRUE, color_palette_20) + theme_bw()

```

# Calculate abundances on resampled rarified data
```{r generate_data_files}
kernel_asv$save_abund(dirpath = "/Volumes/GoogleDrive/My Drive/2022 Wheat MS ASVs/Wheat_seed_endophytes/Data/")
kernel_asv$cal_alphadiv(PD = FALSE)
kernel_asv$save_alphadiv(dirpath = "/Volumes/GoogleDrive/My Drive/2022 Wheat MS ASVs/Wheat_seed_endophytes/Data/")
kernel_asv$cal_betadiv(unifrac = FALSE)
# return dataset$beta_diversity
class(kernel_asv$beta_diversity)
# save dataset$beta_diversity to a directory
kernel_asv$save_betadiv(dirpath = "/Volumes/GoogleDrive/My Drive/2022 Wheat MS ASVs/Wheat_seed_endophytes/Data/")
```

## Merge data for venn plots
```{r vennplots}
venn_merge_site <- kernel_asv$merge_samples(use_group = ("Site"))
venn_plot_site <- trans_venn$new(venn_merge_site, ratio = "seqratio")
p6 <- venn_plot_site$plot_venn()

venn_merge_var <- kernel_asv$merge_samples(use_group = "Variety")
venn_plot_var <- trans_venn$new(venn_merge_var, ratio = "seqratio")
p7 <- venn_plot_var$plot_venn()
```

## Who is different
```{r genus_diff}
genusdiff <- venn_plot_site$trans_comm(use_frequency = TRUE)
# calculate taxa abundance, that is, the frequency
genusdiff$cal_abund()
genusdiff$taxa_abund
# transform and plot
genusdiff2 <- trans_abund$new(dataset = genusdiff, taxrank = "Genus", ntaxa = 8)
p16 <- genusdiff2$plot_bar(bar_type = "part", legend_text_italic = T, ylab_title = "Frequency (%)", xtext_type_hor = FALSE, color_palette_20,
    order_x = c("BB", "BS", "HO", "BB&BS", "BB&HO", "BS&HO", "BB&BS&HO")) +
  theme_bw()
```

# Calculating alpha diversity

## Alpha diversity by site
```{r alphabysite}
alphasite <- trans_alpha$new(dataset = kernel_asv, group = c("Variety","Site"))
alphasite$cal_diff(method = "anova", anova_set = "Site * Variety")

alphasite$res_diff
my_pal<-c("steelblue4", "lightblue", "orange")
p8 <- alphasite$data_alpha %>% 
  filter(Measure == "Observed") %>% 
  ggplot(aes(Variety, Value, color = Site)) +
  geom_boxplot() + 
  ylab("Observed richness") +
  scale_color_manual(values=my_pal) +
  theme_bw() +
  theme(axis.title.y = element_text(size = 20)) +
  theme(axis.title.x = element_text(size = 20))

p8.1 <- alphasite$data_alpha %>% 
  filter(Measure == "Shannon") %>% 
  ggplot(aes(Variety, Value, color = Site)) +
  geom_boxplot() + 
  ylab("Shannon index") +
  scale_color_manual(values=my_pal) +
  theme_bw() +
  theme(axis.title.y = element_text(size = 20)) +
  theme(axis.title.x = element_text(size = 20))
p8.1

```

# Calculating Beta diversity

# Bray-curtis on all samples
```{r brayall}
bray_all <- trans_beta$new(dataset = kernel_asv, group = "Site", measure = "bray")
#set seed can help us reach convergence
set.seed(1)
bray_all$cal_ordination(ordination = "PCoA")
# t1$res_ordination is the ordination result list
class(bray_all$res_ordination)
# plot the PCoA result with confidence ellipse
p9 <- bray_all$plot_ordination(plot_color = "Site", plot_shape = "Variety", plot_type = c("point", "ellipse"), color_palette_20, point_size =5) + theme_bw()
```

## Jaccard for all
```{r brayall}
jacc_all <- trans_beta$new(dataset = kernel_asv, group = "Site", measure = "jaccard")
#set seed can help us reach convergence
set.seed(1)
jacc_all$cal_ordination(ordination = "PCoA")
# t1$res_ordination is the ordination result list
class(jacc_all$res_ordination)
# plot the PCoA result with confidence ellipse
p10 <- jacc_all$plot_ordination(plot_color = "Site", plot_shape = "Variety", plot_type = c("point", "ellipse"), color_palette_20, point_size =5) + theme_bw()
```


```{r}
bray_all$cal_ordination(ordination = "NMDS")
p11 <- bray_all$plot_ordination(plot_color = "Site", plot_shape = "Variety", plot_type = c("point", "ellipse"), color_palette_20, point_size =5) + theme_bw() + xlab("NMDS1") + ylab("NMDS2")
jacc_all$cal_ordination(ordination = "NMDS")
p12 <- jacc_all$plot_ordination(plot_color = "Site", plot_shape = "Variety", plot_type = c("point", "ellipse"), color_palette_20, point_size =5) + theme_bw()

```

## Cluster and PerMANOVA
```{r bray_cluster}
p13 <- bray_all$plot_clustering(group = "Site", replace_name = c("Variety"), color_palette_20)
p14 <- jacc_all$plot_clustering(group = "Site", replace_name = c("Variety"), color_palette_20)
```


```{r}
bray_all$cal_manova(manova_set = "Site * Variety")
bray_all$res_manova
jacc_all$cal_manova(manova_set = "Site * Variety")
jacc_all$res_manova
```

## Beta dispersion with Bray-Curtis
```{r BetadisperBray}
bray_all$cal_betadisper()
bray_all$res_betadisper
```

## Beta dispersion with Jaccard
```{r betadisprJacc}
jacc_all$cal_betadisper()
jacc_all$res_betadisper
```

## Calucualte distances
```{r}
bray_all$cal_group_distance()
bray_all$res_group_distance
p15 <- bray_all$res_group_distance %>% 
  ggplot(aes(Site, value)) +
  theme_bw() +
  geom_boxplot() +
    scale_x_discrete(labels = c("Samples", "BB", "BS", "HO"), limits = c("Samples", "BB", "BS", "HO")) +
  geom_boxplot(aes(value), position = position_nudge( x = 0.475)) +
  ylab("Bray-Curtis distance") +
  xlab("") +
  
  theme(axis.title.y = element_text(size = 20)) +
  theme(axis.text.x = element_text(size = 18))
  
```

## Calculate taxa abundance
```{r}
raw_taxa<- as.data.frame(t(kernel_asv$otu_table))
raw_taxa <- raw_taxa/rowSums(raw_taxa)
wheatRA <- cbind(kernel_asv$sample_table,raw_taxa)
taxa_wheat <- tax_tab %>%
  rownames_to_column(var="ASV")
wheatRA <- wheatRA %>% 
  select(-SampleID) %>% 
  rownames_to_column(var = "SampleID") %>% 
  select(-Group, -Rep) %>% 
  pivot_longer(4:90,
               names_to = "ASV",
               values_to = "Abundance") %>% 
  left_join(taxa_wheat) %>% 
  view()
```

## Make a taxa plot function
```{r taxapllotFUN}
taxa_plot <- function(x,y,z){
  ggplot(x, aes(x = Variety, y = Abundance, shape=Site)) + 
    ggtitle(z) +
  geom_jitter(size = 3, color = y) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        panel.border=element_rect(colour = "black", fill=NA))+
  ylab("Raw Abundance")+
  theme(plot.title=element_text(hjust = 0.5, size = 22),
        axis.text=element_text(size = 16),
        axis.title=element_text(size = 20),
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1)) +
  geom_vline(xintercept = 1.5, linetype="dashed", 
             color = "grey", size=1) +
  geom_vline(xintercept = 2.5, linetype="dashed", 
             color = "grey", size=1)
}
```

#Plotting specific taxa 
```{r}
PAN<-wheatRA %>% 
  filter(Genus == "Pantoea")
p17 <- taxa_plot(PAN,"red", "Pantoea")

PSE<-wheatRA %>% 
  filter(Genus == "Pseudomonas")
p18 <- taxa_plot(PSE,"purple", "Pseudomonas")

LAC<-wheatRA %>% 
  filter(Genus == "Lactobacillus")
p19 <- taxa_plot(LAC, "darkorange4", "Lactobacillus")

MET<-wheatRA %>% 
  filter(Genus == "Methylobacterium")
p20 <- taxa_plot(MET, "deepskyblue4", "Methylobacterium") 

p21 <- arrangeGrob(p17,p18,p19,p20, nrow = 2, ncol = 2)
ggsave("taxa_panel.png", p21)
```


```{r}
pathos <- wheatRA %>% 
  filter(Genus == "Curtobacterium") %>%
  filter(Abundance > 0) %>% 
  ggplot(aes(Site, Abundance, color = ASV)) +
  geom_boxplot()

curt_asvs <- wheatRA %>% 
  filter(Genus == c("Curtobacterium")) %>%
  group_by(ASV) %>% 
  summarise(mean = mean(Abundance)*100, n = n()) %>% 
  view()
  
```

## Generate a plot with RA by site faceted by Variety
```{r}
colourCount<- length(unique(WheatRA2_long2$Family))
getPalette<- colorRampPalette(brewer.pal(12, "Paired"))
ra_site_genus<-ggplot(wheatRA, aes(x = Site, y = Abundance, fill = Genus)) + 
  geom_bar(stat = "identity", position="fill") +
  scale_fill_manual(values = getPalette(colourCount)) +
  facet_wrap(vars(Variety)) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        panel.border=element_rect(colour = "black", fill=NA)) +
  ylab("Relative Abundance") +
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=26),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust =1))
ra_site_genus
```

## Generate a plot with Family RA by site faceted by Variety
```{r}
ra_site_fam<-ggplot(wheatRA, aes(x = Site, y = Abundance, fill = Family)) + 
  geom_bar(stat = "identity", position="fill") +
  scale_fill_manual(values = getPalette(colourCount)) +
  facet_wrap(vars(Variety)) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        panel.border=element_rect(colour = "black", fill=NA)) +
  ylab("Relative Abundance") +
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=26),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust =1))
ra_site_fam
```

## Generate a plot with Family RA by site faceted by Variety
```{r}
ra_site_phy<-ggplot(wheatRA, aes(x = Site, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", position="fill") +
  scale_fill_manual(values = getPalette(colourCount)) +
  facet_wrap(vars(Variety)) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        panel.border=element_rect(colour = "black", fill=NA)) +
  ylab("Relative Abundance") +
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=26),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust =1))
ra_site_phy
```
